Homicide Increases in the COVID-19 Era

### 2025-02-28 Draft2.qmd
Creating Draft2 version to submit

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions and Crimes in the City of Los Angeles \nFebruary 2023 to February 2025", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_stata(scheme = "s1mono") +
  theme(axis.text.x = element_text(angle = 300, vjust = 0.5, hjust = -.1),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```
### Interpretation
This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data. By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions. Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month. There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average. Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.

@@

# covid data
covid_14day$date <- ymd(covid_la$ep_date)
covid_14day$month <- month(covid_la$date)
covid_14day$year <- year(covid_la$date)
covid_14day$date <- day(covid_la$date)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_14day$lagged <- covid_la$date + 7

@@

covid_la_monthly$index <- 1:nrow(covid_la_monthly)
# base plot shows that this does in fact mirror LA county so the case numbers should be accurate
# plot(y = covid_la_monthly$cases, x = covid_la_monthly$index, type = "l")

crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
crime_la_monthly$month <- month(crime_la_monthly$date)
crime_la_monthly$year <- year(crime_la_monthly$date)
```

>>>>

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions and Crimes in the City of Los Angeles \nFebruary 2023 to February 2025", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_stata(scheme = "s1mono") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```
### Interpretation
This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data. By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions. Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month. There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average. Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.

@@

# covid data
covid_14day$date <- ymd(covid_14day$ep_date)
covid_14day$month <- month(covid_14day$date)
covid_14day$year <- year(covid_14day$date)
covid_14day$date <- day(covid_14day$date)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_14day$lagged <- covid_14day$date + 7

@@

covid_la_monthly$index <- 1:nrow(covid_la_monthly)

# next, unemployment rate
unemployment <- read.csv("Raw_Data/la_county_unemployment_rate.csv")
unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))

# lapd budget
lapd_budget <- read.csv("Raw_Data/lapd_budget.csv")

crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
crime_la_monthly$month <- month(crime_la_monthly$date)
crime_la_monthly$year <- year(crime_la_monthly$date)

correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")

correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))

correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")

correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")

correlations <- correlations_df4 %>% select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)

Cor_mat <- round(cor(correlations), 2)

long_Cor <- reshape2::melt(Cor_mat)

long_Cor <- long_Cor %>%
    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) + 
  ggtitle("Correlations Between the Number of Crimes and \n Different Law Enforcement and Socioeconomic Factors")
```
